---
title: "Excercise 1"
author: "Nam Tran"
output:
  md_document: 
     variant: markdown_github
  html_document: default
---
## Part 1

In this part, I aimed to find a relationship that could dispel the notion that green ratings equate to higher rents, and thus higher revenues.I first started by grouping the buildings by whether or not they attained a green rating.  Here we see some quick differences in rent, leasing rate etc.

```{r setup, include= FALSE}
library(plyr)
library(tidyverse)
library(FNN)
library(gridExtra)
greenbuildings=read.csv('../Data Files/greenbuildings.csv')
ABIA = read.csv('../Data Files/ABIA.csv')
```

```{r, echo=FALSE}
g1 <- ddply(greenbuildings, "green_rating", summarise, rent.mean=mean(Rent), leasing.mean=mean(leasing_rate), gas.mean=mean(Gas_Costs), electric.mean=mean(Electricity_Costs), age.mean=mean(age))
c3 = c("A","B")
g1 = cbind(g1,c3)
bar1 <- ggplot(g1, aes(green_rating, rent.mean,fill = c3)) +
  geom_bar(stat='identity',width= .75) +
  scale_x_continuous(breaks = seq(0,1)) +
  scale_fill_manual("legend", values = c("A" = "gray27", "B" = "forestgreen")) +
  labs(
    title = "Higher rent in green buildings",
    x = "Green Status",
    y = "Rent"
  ) +
  theme_classic()+
  theme(axis.title.x=element_blank())+
  theme(legend.position = "none")


bar2 <- ggplot(g1, aes(green_rating,leasing.mean, fill = c3)) +
  geom_bar(stat='identity',width= .75) +
  scale_x_continuous(breaks = seq(0,1)) +
  scale_fill_manual("legend", values = c("A" = "gray27", "B" = "forestgreen")) +
  labs(
    title = "Higher leasing rates in green buildings",
    x = "Green Status",
    y = "Leasing Rates"
  ) +
  theme_classic()+
  theme(axis.title.x=element_blank())+
  theme(legend.position = "none")


bar3 <- ggplot(g1, aes(green_rating, gas.mean, fill = c3)) +
  geom_bar(stat='identity',width= .75) +
  scale_x_continuous(breaks = seq(0,1)) +
  scale_fill_manual("legend", values = c("A" = "gray27", "B" = "forestgreen")) +
  labs(
    title = "Lower gas costs in green buildings",
    x = "Green Status",
    y = "Gas Costs" 
  ) +
  theme_classic()+
  theme(axis.title.x=element_blank())+
  theme(legend.position = "none")


bar4 <-ggplot(g1, aes(green_rating, electric.mean, fill= c3)) +
  geom_bar(stat='identity',width= .75 ,aes(fill = )) +
  scale_x_continuous(breaks = seq(0,1)) +
  scale_fill_manual("legend", values = c("A" = "gray27", "B" = "forestgreen")) +
  labs(
    title = "Higher electricity costs in green buildings",
    y = "Electricty Costs"
  ) +
  theme_classic()+
  theme(axis.title.x=element_blank())+
  theme(legend.position = "none")

grid.arrange(bar1,bar2,bar4,bar3, ncol=2, nrow=2) #first set of graphs

```


The pros of green buildings are evident here, but are the higher rents due to green ratings or other variables? Next, I examined the relationship between rent and age.


```{r, echo= FALSE, message=FALSE}
greenbuildings <- mutate(greenbuildings,age.quantile = ntile(age,10))
greenbuildings$age.quantile = factor(greenbuildings$age.quantile)
g2 <- ddply(greenbuildings, "age.quantile", summarise, age.rent=mean(Rent), age.leasing=mean(leasing_rate))
age.bar <- ggplot(g2, aes(age.quantile, age.rent )) +
  geom_bar(stat='identity',width=.5) +
  theme_light()+
  labs(
    title = "Decreasing rent as age increases",
    y = "Rent",
    x = "Quantile"
  )
  

age.scatter <- ggplot(greenbuildings, aes(age,Rent))+
  geom_point(aes(color = age.quantile)) +
  geom_smooth(se = FALSE,color= "red") +
  scale_color_hue(l=55, c=70) +
  theme_light()+
  theme(legend.position = "bottom")+
  labs(
    title= "",
    y = "Rent",
    x = "Age",
    color='Quantile'
  )

greenbuildings = greenbuildings %>%
  mutate(green.factor = factor(greenbuildings$green_rating))

g3 <- ddply(greenbuildings, "green.factor", summarise, age.mean=mean(age),age.median= median(age))

age.density <- ggplot(greenbuildings, aes(x=age, fill = green.factor)) +
  geom_density(alpha=.4)+
  scale_fill_manual(values=c("#999999", "forestgreen", "#56B4E9"))+
  scale_color_manual(values=c("#999999", "forestgreen", "#56B4E9")) +
  geom_vline(data=g3, aes(xintercept=age.median, color=green.factor),
             linetype="dashed")+
  labs(
    title= "Higher distribution of green ratings in lower age ranges",
    y = "Density",
    x = "Age",
    fill ='Green Rating',
    color = 'Median Age'
  )+
  theme_light()

age.grid = grid.arrange(age.bar,age.scatter, ncol=2, nrow=1) #second set of graphs
age.density

```

Here, it is clear that rent is lower in the newer buildings. I did this by dividing the age into ten quantiles and finding the average rent in each quantile. The age density graph serves to reinforce the notion that green buildings are much newer than non-green buildings. This certainly doesn't prove much, but it does make the "data guru"'s assertions a bit odd.

## Part 2

In this section, I mainly looked to answer two questions: What is the best time of day to fly to minimize delays and What is the best time of year to fly to minimize delays?
Here are the results:
```{r, echo = FALSE, message= FALSE, warning= FALSE}
ABIA= ABIA %>%
  mutate(TimeofDay=
           ifelse(between(ABIA$CRSDepTime,0,600), "Early Morning", 
                  ifelse(between(ABIA$CRSDepTime,601,1200),"Morning" , 
                         ifelse(between(ABIA$CRSDepTime,1201,1800),"Afternoon",
                                ifelse(between(ABIA$CRSDepTime,1801,2400),"Night",0)
                        )
                  )
           )
  )
ABIA$TimeofDay= factor(ABIA$TimeofDay)
g4 <- ddply(ABIA, "TimeofDay", summarise, DelayTime.mean=mean(na.exclude(DepDelay)))


Time.Delay <-ggplot(g4, aes(x = reorder(TimeofDay, -DelayTime.mean), y = DelayTime.mean))+
  geom_bar(stat='identity',width=.5,fill= "lightpink", colour="black") +
  scale_y_continuous(breaks = seq(-2,24,2))+
  labs(
    title = "Increasing delay times throughout the day",
    y = "Departure Delay in Minutes"
  )+
  theme_bw()+
  theme(axis.title.y=element_blank())+
  coord_flip()

ABIA= ABIA %>%
  mutate(Season=
           ifelse(between(ABIA$Month,3,5), "Spring", 
                  ifelse(between(ABIA$Month,6,8),"Summer" , 
                         ifelse(between(ABIA$Month,9,11),"Autumn",
                                ifelse(between(ABIA$Month,1,2)|ABIA$Month == 12,"Winter",0)
                         )
                  )
           )
  )
g5 <- ddply(ABIA, "Season", summarise, DelaySeason.mean=mean(na.exclude(DepDelay)), DelaySeason.median=median(na.exclude(DepDelay)))

Season.Delay <- ggplot(g5, aes(x = reorder(Season, -DelaySeason.mean), y = DelaySeason.mean))+
  geom_bar(stat='identity',width=.5,fill= "tomato1", colour="black") +
  scale_y_continuous(breaks = seq(-2,24,2))+
  labs(
    title = "Less delay times during Autumn",
    y = "Departure Delay in Minutes"
  )+
  theme_bw()+
  theme(axis.title.y=element_blank())+
  coord_flip()
Time.Delay
```

For the times of day, I divided the 24 hours into 4, 6 hour periods. Early morning flights ran from 12 a.m. to 6 a.m., morning flights from 6.a.m. to 12 p.m. and onwards. Unsurprisingly, the least amount of delays occured in the "early morning" time frame, and in fact, the average flight was ahead of time! When looking at seasons, the results are also rather unsurprisng.

```{r, echo = FALSE, message= FALSE, warning= FALSE}
Season.Dist <- ggplot(ABIA, aes(x=DepDelay,fill= Season)) +
  geom_density(alpha=.6)+
  scale_fill_manual(values=c("darkorange2", "green3", "yellow", "ivory"))+
  facet_wrap(~Season,nrow = 2)+
  geom_vline(data=g5, aes(xintercept= DelaySeason.mean, color ="Mean"),
             linetype="F1") +
  scale_color_manual(values=c("grey22"))+
  scale_x_continuous(breaks = seq(0,900, 100)) +
  labs(
    title = "Seasonal departure delay distribution",
    y = "Density",
    x = "Departure Delay in Minutes"
  )+
  theme_bw() +
  guides(fill=FALSE)+
  theme(legend.title=element_blank())
Season.Delay
Season.Dist

```

For the seasons, I divided the months by the traditional definitions. Winter runs from December to february, Spring going from March to May, Summer spanning June through August, and Autumn rounding it out from September to November. These results follow traditional intuition. Autumn doesn't span many major holidays, and it is also one of the more fair weather seasons in this area. I quickly also looked at delay times by destination. The only major surprise is the absurdly high delay time from DSM which is Des Moines International Airport. Perhaps it is just a victim of severe outliers due to extreme weather or other unlikely delays.

```{r, echo = FALSE, message= FALSE, warning= FALSE}
g6 <- ddply(ABIA, "Dest", summarise, DelayDest.mean=mean(na.exclude(DepDelay)))
g6 <- na.omit(g6)
Dest.Delay <- ggplot(g6, aes(x = reorder(Dest,DelayDest.mean), y = DelayDest.mean))+
  geom_bar(stat='identity',width=.5,fill= "tomato1", colour="black") +
  labs(
    title = "Worst delay by destination",
    y = "Departure Delay in Minutes"
  )+
  scale_y_continuous(breaks = seq(-10,200,10))+
  theme_grey()+
  theme(axis.title.y=element_blank())+
  theme(text = element_text(size=6),axis.text.y = element_text(angle = 35,hjust = 1,vjust= 0))+
  coord_flip()
Dest.Delay
```

## Part 3

In this part, I begin by dividng the data by two specific trims, 350 and 65 AMG. Then after creating the training and testing splits, I use K-nearest neighbors to build a model. I run each trim from K = 3 to K = 100, finding the out of sample rmse for each. I then used the minimum rsme of the set to build the final predictive graph.

```{r, echo = FALSE, message= FALSE, warning= FALSE}

sclass = read.csv('../Data Files/sclass.csv')

sclass350 = subset(sclass, trim == '350')

sclass65AMG = subset(sclass, trim == '65 AMG')

N = nrow(sclass350)
N_train = floor(0.8*N)
N_test = N - N_train

N350 = nrow(sclass350)
N_train350 = floor(0.8*N350)
N_test350 = N350 - N_train350

train_ind350 = sample.int(N350, N_train350, replace=FALSE)


D_train350 = sclass350[train_ind350,]
D_test350 = sclass350[-train_ind350,]

D_test350 = arrange(D_test350, mileage)

X_train = select(D_train350, mileage)
y_train = select(D_train350, price)
X_test = select(D_test350, mileage)
y_test = select(D_test350, price)

rmse = function(y, ypred) {
  sqrt(mean(data.matrix((y-ypred)^2)))
}

variables =1
iterations = 100
RMSE350 <- matrix(ncol=variables, nrow=iterations)

for (i in 3:iterations)
{
  knn = knn.reg(train = X_train, test = X_test, y = y_train, k=i)
  ypred_knn = knn$pred
  rmse1 <- rmse(y_test, ypred_knn)
  RMSE350[i,]<-rmse1
}

RMSE350 <-data.frame(RMSE350)
RMSE350<- mutate(RMSE350, k = (1:100))
RMSE350 <- RMSE350[-c(1,2),]
colnames(RMSE350)[colnames(RMSE350)=="RMSE350"] <- "rmse"


N = nrow(sclass65AMG)
N_train = floor(0.8*N)
N_test = N350 - N_train


train_ind = sample.int(N, N_train, replace=FALSE)

D_train = sclass65AMG[train_ind,]
D_test = sclass65AMG[-train_ind,]

D_test = arrange(D_test, mileage)

X_train1 = select(D_train, mileage)
y_train1 = select(D_train, price)
X_test1 = select(D_test, mileage)
y_test1 = select(D_test, price)

variables =1
iterations = 100
RMSE65AMG <- matrix(ncol=variables, nrow=iterations)

for (i in 3:iterations)
{
  knn = knn.reg(train = X_train1, test = X_test1, y = y_train1, k=i)
  ypred_knn = knn$pred
  rmse2 <- rmse(y_test1, ypred_knn)
  RMSE65AMG[i,]<-rmse2
}

RMSE65AMG <-data.frame(RMSE65AMG)
RMSE65AMG <- mutate(RMSE65AMG, k = (1:100))
RMSE65AMG <- RMSE65AMG[-c(1,2),]
colnames(RMSE65AMG)[colnames(RMSE65AMG)=="RMSE65AMG"] <- "rmse"

rmsek65AMG <- ggplot(RMSE65AMG, aes(k , rmse))+
  geom_path(aes(x= k, y = rmse))+
  labs(
    title =  "Rmse for 65 AMG trim"
  )

rmsek350 <- ggplot(RMSE350, aes(k , rmse))+
  geom_path(aes(x= k, y = rmse))+
  labs(
    title = "Rmse for 350 trim"
  )
grid.arrange(rmsek65AMG,rmsek350, ncol=2)

cat("Optimal K for 65 AMG trim: ", (which(RMSE65AMG==min(RMSE65AMG$rmse))+2))

cat("Optimal K for 350 trim: ", (which(RMSE350==min(RMSE350$rmse))+2))

knnamg = knn.reg(train = X_train1, test = X_test1, y = y_train1, k=(which(RMSE65AMG==min(RMSE65AMG$rmse))+2))
ypred_knnamg = knnamg$pred
D_test$ypred_knnamg = ypred_knnamg

p_test = ggplot(data = D_test) + 
  geom_point(mapping = aes(x = mileage, y = price), color='lightgrey')+
  geom_path(aes(x = mileage, y = ypred_knnamg), color='red') +
  labs(
    title = "Prediction for 65 AMG"
  )+
  theme_bw()
p_test


knn350= knn.reg(train = X_train, test = X_test, y = y_train, k=(which(RMSE350==min(RMSE350$rmse))+2))
ypred_knn350 = knn350$pred
D_test350$ypred_knn350 = ypred_knn350

p_test1 = ggplot(data = D_test350) + 
  geom_point(mapping = aes(x = mileage, y = price), color='lightgrey')+
  geom_path(aes(x = mileage, y = ypred_knn350), color='red') +
  labs(
    title = "Prediction for 350"
  )+
  theme_bw()
p_test1

```

Overall, the 350 trim  has a higher optimal K value. I think this is due to the larger sample size available. It enables the model to "explain" more of the variance. However, with this training and test split, the results do vary from run to run.
