---
title: "Examining the Evolution of 3-point Attempts"
author: "Nam Tran"
date: "May 16, 2019"
output: 
  md_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(tidyverse)
library(ggrepel)
library(ggalt)
library(stringr)
library(rpart)
library(gbm)


playerdata = read.csv('../Data Files/Seasons_Stats.csv')
playerbio = read.csv('../Data Files/player_data.csv')
names(playerbio)[1]<-"Player"

players<- left_join(playerdata, playerbio)
birthyear <- str_split_fixed(players$birth_date, ", ", 2)
players$birthyear <- birthyear[,2]
players$birth_date <- NULL


playerdata  <- subset(playerdata, Year>= 1980)

clplayerdata <- playerdata %>% 
  distinct(Player,Year, .keep_all = TRUE)
clplayerdata <-subset(clplayerdata, X3PA >= 80)
clplayerdata[3:31] <-NULL
clplayerdata[c("X","blank2","blanl","year_start", "year_end", "position", "height","DRB", "ORB","TRB", "college","X3P","X2P","FT")] <-NULL


pc_player = prcomp(clplayerdata,rank = 5 , scale= TRUE)
pc_player
summary(pc_player)
clplayerdata = cbind(clplayerdata,pc_player$x)

clplayerdata1 <- playerdata[!playerdata$Tm %in% c("TOT"),]

clplayerdata2 <- playerdata %>% 
  distinct(Player,Year, .keep_all = TRUE)



X3PA.year <- ddply(clplayerdata2, "Year", summarise, sum=sum(X3PA), mean=mean(X3PA))
X3PA.year = X3PA.year  %>%
  mutate(Exp =
           ifelse(X3PA.year$Year<=1995, 0, 
                  ifelse(X3PA.year$Year > 1995 & X3PA.year$Year <= 2004, 1, 2
                         )
                  )
)
X3PA.year$Exp <- factor(X3PA.year$Exp)

X3PA.Tm <- ddply(clplayerdata1, .(Tm,Year), summarise, sum=sum(X3PA), mean=mean(X3PA))

X3PA.Tm = X3PA.Tm %>%
  mutate(X3PAG = sum/82)
X3PA.Tm2016 <- subset(X3PA.Tm, Year == 2016)
X3PA.Tm2017 <- subset(X3PA.Tm, Year == 2017)
X3PAG <- data.frame("Tm"= X3PA.Tm2017$Tm, X2016= X3PA.Tm2016$X3PAG, X2017= X3PA.Tm2017$X3PAG)
median(X3PAG$X2017)
X3PAG <-subset(X3PAG, X2016 >=median(X3PAG$X2016))
X3PAG =X3PAG %>%
  mutate(difference = X2017-X2016)
X3PAG$Tm <- factor(X3PAG$Tm, levels = X3PAG$Tm[order(X3PAG$difference)])


Min3PA2017 <- subset(clplayerdata2, X3PA >= 100 & Year ==2017)
Min3PA2017 = Min3PA2017  %>%
  mutate(X3PS =
           ifelse(Min3PA2017$X3P.<=.1, "terrible", 
                  ifelse(Min3PA2017$X3P.<=.2 & Min3PA2017$X3P. >= .1,"poor" , 
                         ifelse(Min3PA2017$X3P.<=.3 & Min3PA2017$X3P. >= .2,"average",
                                ifelse(Min3PA2017$X3P.<=.4 & Min3PA2017$X3P. >= .3,"good","excellent")
                         )
                  )
           )
  )


# Plot 1
Total3PA<- ggplot(X3PA.year, aes(x=Year, y=sum)) + 
  geom_point(size=3) + 
  geom_vline(linetype= "twodash",xintercept =c(1988,1989,1995,2004), color ='dark red')+
  geom_segment(aes(x=Year, 
                   xend=Year, 
                   y=0,
                   yend=sum)) + 
  labs(title="Figure 1",
       subtitle = "Increasing Three Point Attempts since 1980",
       x=NULL,
       y="Total 3PA",
       caption ="*Dashed lines denote expansion years in which 1 or 2 teams join the NBA") + 
  theme(axis.text.x = element_text(angle=35, vjust=0.6)) +
  theme(axis.text.y = element_text(angle=35, vjust = .05, hjust=1)) +
  theme(plot.caption = element_text(face ="italic"))+
  scale_x_continuous(breaks = round(seq(min(X3PA.year$Year), max(X3PA.year$Year), by = 5),1))+
  scale_y_continuous(breaks = round(seq(0, max(X3PA.year$sum), by = 10000),1)) 

Total3PA


gg1 <- ggplot(Min3PA2017, aes(X2PA,X3PA, label = Player)) + 
  geom_jitter(aes(color=X3PS)) +
  geom_label_repel(data          = subset(Min3PA2017, X3PA > 646),
                  size          = 3,
                  box.padding   = .5,
                  point.padding = 0.5,
                  force         = 1,
                  segment.size  = 0.2,
                  segment.color = "grey50",
                  direction     = "both") +
  geom_label_repel(data          = subset(Min3PA2017, X2PA > 1300),
                   size          = 3,
                   box.padding   = .5,
                   point.padding = 1,
                   force         = 1,
                   segment.size  = 0.2,
                   segment.color = "grey50",
                   direction     = "y") +
  labs(title="Figure Two",
       subtitle= "Shot Selection in the 2016-2017 season",
       x="2-Point FG Attempts",
       y="3-Point FG Attempts",
       caption ="Minimum 100 3-point FG attempts") +
  scale_x_continuous(breaks = round(seq(0, max(Min3PA2017$X2PA), by = 200),1)) +
  scale_colour_manual(name = '3-point %', 
                      values =c("orangered", "deepskyblue2", "forestgreen"), labels = c('20-30%','30-40%','40% +'))

gg1

gg2 <- ggplot(X3PAG, aes(x=X2016, xend=X2017, y=Tm)) + 
  geom_dumbbell(color="black", 
                size_x=1.5, 
                size_xend = 2.25,
                colour_x="lightblue3", 
                colour_xend = "#0e668b",
                show.legend =TRUE)+
  labs(x="3-point Attempts Per Game", y=NULL, 
       title="3-Point Attempts Per Game", 
       subtitle=" From Change: 2016 (light blue) to 2017 (dark blue)")+
  scale_colour_manual(name = '3-point %', 
                      values =c("orangered", "deepskyblue2"), labels = c('20-30%','30-40%'))

gg2

N= nrow(clplayerdata)

train_frac=.8
N_train= floor(train_frac*N)
N_test =N-N_train
train_ind=sample.int(N,N_train, replace = FALSE) %>% sort
player_train = clplayerdata[train_ind,]
player_test =clplayerdata[-train_ind,]

big.tree = rpart(X3PA ~.,method="anova",data=player_train,control=rpart.control(minsplot=5, cp=1e-6,xval=10))

nbig = length(unique(big.tree$where))

plotcp(big.tree)

head(big.tree$cptable)

err_1se = big.tree$cptable[,'xerror'] + big.tree$cptable[,'xstd']
err_thresh = min(err_1se)
err_thresh
big.tree$cptable[,'xerror'] - err_thresh
which(big.tree$cptable[,'xerror']- err_thresh <0) %>% head

bestcp= big.tree$cptable[166,'CP']
bestcp

cvtree = prune(big.tree,cp=bestcp)
plot(cvtree)
length(unique(cvtree$where))
plot(player_train$Year,predict(cvtree))
plot(player_train$X3PA,predict(cvtree))


yhat_test_tree = predict(cvtree,player_test)

rmse = mean((yhat_test_tree - player_test$X3PA)^2) %>% sqrt
rmse
```


# Abstract
In this report, I summarize the overall trends of the 3-point attempts starting from 1980 to 2017. With the 3-point shot being introduced in 1979, we have seen an increase in volume from 1980 to 2017 at astounding rates, even with the multiple expansion years. In the most recent season of the data set ,2017, we see four out of the five prolific shooters, being on two of the winningest teams in the league, GSW being first and HOU placing third. From our predicitve model, I got an astoundingly low RMSE at **47.73**. The model predicts linearly with the training data set, which is a positive sign for its overall predicitive powers. Among the predictions, the 3-point attempts can be seen rising steadily over the years, and this trend will likely continue with the likes of James Harden, Klay Thompson, and Steph Curry at the forefront.


# Introduction
In today's age of basketball, we've all heard of the names of prolific 3-point shooters such as Steph Curry, Klay Thompson, and James Harden, and the wave of 3-point shooting that has accompanied them. Harden in particular has been indoctrinated into a system revolving around this shot. The Houston Rockets, a home town favorite of mine, are notorious around the league for pushing the envelope of this new 3-point movement, in which  they have dropped the midrange 2-point shot from their repertoire in favor of the more efficient 3-point shot.

Teams have not alwasy been so reliant on this particular shot or even had the opportunity for it. Surprising to many, the 3-point shot was only introduced to the NBA in the 1979 season, and other governing basketball bodies such as FIBA, Olypmics, and NCAA followed suit in the next few years. The NBA since then has tried to tinker with the distance of the shot to increase and decrease scoring, but today it is at the original 23 feet and 9 inches around the arc while being 22 ft. from the corners.

But where will all this lead? Has the likes of Steph Curry and James Harden ushered in a new era of basketball dependent on the 3-point shot? Here we will examine this increase in volume of 3-point shots using other advanced basketball metrics.

# Methodology

The data set used here was obtained through Kaggle, but it is a aggregated dataset of players from 1950-2017 scraped from  basketball-reference.com. The data includes the typical box-score stats such as points, assists, and rebounds to advanced features such as PER and true shooting percentages.

To examine the trends using data visualizations, I began by mutating the data and finding summary variables typically by year or team. I wanted to look at the overall trend from the 80's until 2017 as well as 2017 by itself. It's important to note how the league first started with the 3-point shot, and how players are choosing their shots in the modern era.

For  my prediction model, I begin by subsetting the data by years after 1979 where the NBA introduced the 3-point line. From their I look into other correlation and trends that may be interesting. Here, I start  with teams and years and then move towards basketball metrics. The main statistic that I am looking at is X3PA or 3-point attempts. I wanted a baseline of some of the more "proflific" shooters, so i subsetted the data once again by X3PA >= 100. This removes a majority of bench or role players that may put up only a few shots a season. After subsetting twice, I end up with 4432 observations.

I wanted to use something we recently learned to be quite powerful, **trees**. For this purpose, I removed all of the advance basketball metrics like PER amd TS%. I believe while these are interesting to look at they would be confounding in this model.I aslo looked to remove 3-points made, 2-points made, and free throws made, while keeping attempts and percentages for each of these.
I then use Principle Components Analysis (PCA) to generate five additional variables that will help the tree find the proper nodes. With these componenets, I have a total of 22 variables to run the tree on.

Now for our cleaned dataset, I will run the tree, find the best complexity parameter, and prune the tree using this complexity parameter.


# Results and Analysis
The basic trend noted earlier that 3-point shot volume has increased in the past 3 decades is clearly seen in Figure 1. From the introduction of the shot to 2017, volume has increased by  nearly 12 times. It is to be noted that, in 1980 there were only 23 teams, compared to the modern era of 30 teams. Despite the difference in number of teams, this pace is still astounding. the largest. The three season spike in the 90's was likely due to efforts by the NBA to increase scoring by decreasing the 3-point distance. In the following year, it was returned to its original distance, and the 3-point trend continues to climb up.

```{r echo= FALSE}
Total3PA
```

Next we look turn to look at the most recent year in Figure Two. When we look at the highlighted names towards the top of 3-point FG attempts, we see that four out of five of these players are on the **same** team. What also sticks out is a player like Demar Derozan, a shooting guard, leading this group in 2-point FG attempts. In this day and age, it'd almost be absurd to take this many two's at his position. 

```{r echo= FALSE}
gg1
```

Moving on to the predictive model portion, after generating a big tree, and pruning it down. The model RMSE is at an exceptionally low 47.73. I'm not too sure why it's so low, perhaps the additions of PCA really help drive the tree and it's predictive powers.

